---
title: "Homework 5"
author: "Jeff Kropelnicki"
date: '`r format(Sys.Date())`'
output:
  html_document: default
  pdf_document: default
---   
   
      
```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(knitr)
library(dplyr)
library(kableExtra)
library(readxl)
library(e1071)
```   
   
```{r, message=FALSE, warning=FALSE, echo=FALSE}
data <- read_excel("CellDNA.xls", col_names = FALSE)
```   
   
      
Standerize the cell data and make the last column a 0 or 1. If a number is greater than 0 make it a 1.    
```{r}
#Make any number is the last column a 0 or 1
data$X__14[ data$X__14 > 0 ] = 1

#Normalize the data is the X columns 1-13
norm <- data[,1:13]
x <- as.matrix(scale(norm))
y <- as.matrix(data[,14])

#Take a look at the data
quick_look <- as.data.frame(cbind(x,y))

kable(head(quick_look), "html") %>%
          kable_styling(bootstrap_options = "striped", font_size = 10)
```     
   
      
```{r}
#Build and run the SVM model with the e1071 package. This is a linear classification SVM model with a cost of 10. Cost is the cost of miss classifications.  
set.seed(777)
model <- svm(factor(X__14) ~ ., data = quick_look, type='C-classification', kernel='linear',scale=FALSE, cost = 10) 
summary(model)
```   
**Question 1**   
There are `r model$tot.nSV` support vectors. There are `r model$nSV[1]` in the 0 class and `r model$nSV[2]` in the 1 class.
   
      
```{r}
#The weight overall is 
w <- as.vector(t(model$coefs) %*% model$SV)
b <- -model$rho
p <- model$SV
```   
   
**Question 2:** List top 3 records that have the smallest **absolute** values from $w^TX + b$.      
```{r}
#Return the co
t1 <- as.data.frame(abs(model$coefs)) %>%
      arrange(V1)
kable(head(t1, n = 3), "html") %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```   
   
      
```{r}
#Do a prediction for 131, 165, 892, 1057 
newdata = quick_look[c(131, 165, 892, 1057), ]
predict(model, newdata)
pred <- predict(model, newdata)

table(Actual=newdata$X__14, Fitted=pred)

#Theta for 131
z131 <- as.vector(sum(t(quick_look[131, -14])*w)+b)

#Theta for 165
z165 <- as.vector(sum(t(quick_look[165, -14])*w)+b)

#Theta for 892
z892 <- as.vector(sum(t(quick_look[892, -14])*w)+b)

#Theta for 1057
z1057 <- as.vector(sum(t(quick_look[1057, -14])*w)+b)
```   
   
      
**Question 3:**   
The $w^TX + b$” values for the record 131 is `r z131`   
The $w^TX + b$” values for the record 165 is `r z165`   
The $w^TX + b$” values for the record 892 is `r z892`   
The $w^TX + b$” values for the record 1057 is `r z1057`    
The 0 varibles 131 and 1057 are on the positive side of the model and 1 varibles 165 and 892 are on the negitive side of the model. 




















