---
author: "Jeff Kropelnicki"
title: "Homework 4"
date: "`r format(Sys.Date())`"
output:
  html_document: default
  pdf_document: default
---   
   
      
```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(knitr)
library(dplyr)
library(tidyr)
library(kableExtra)
library(readxl)
library(glmnet)
```   
   
      
Import the cell DNA data  
```{r, message=FALSE, warning=FALSE}
data <- read_excel("CellDNA.xls", col_names = FALSE)
```   
   

Standerize the cell data and make the last column a 0 or 1. If a number is greater than 0 make it a 1.    
```{r}
#Make any number is the last column a 0 or 1
data$X__14[ data$X__14 > 0 ] = 1

#Normalize the data is the X columns 1-13
norm <- data[,1:13]
x <- as.matrix(scale(norm))
y <- as.matrix(data[,14])

#Take a look at the data
quick_look <- as.data.frame(cbind(x,y))

kable(head(quick_look), "html") %>%
          kable_styling(bootstrap_options = "striped", font_size = 10)
```   
   
      
```{r}
#Build the lasso L1 model
set.seed(999)
glmmod <- cv.glmnet(x, y, alpha=1, nfolds = 10, family="binomial")


par(mfrow=c(1,2))
plot(glmmod)
plot(glmmod$glmnet.fit, xvar="lambda", label = TRUE)
```   
   
Lets see what the lambdas are for the min and 1 standard deviation from the min are 
```{r}
#The min 
t1 <- round(glmmod$lambda.min, 6)
#1 SD from the min
t2 <- round(glmmod$lambda.1se, 8)
```   

The minimum lambda is `r t1` and the lambda that is one standard deviation from the minimum is `r t2`   
   
```{r}
#predictors at mim lambda
coef(glmmod, s=glmmod$lambda.min)
```   
At the minimum lambda 0.00024 there are 11 predictors. 
   
      
         
Lets see what predictors are selected with lambda this is one standard deviation from the minimum.  
```{r}
#predictors at 1 sd from min lambda
coef(glmmod, s=glmmod$lambda.1se)
```   
At the lambda 0.0068 that is one standard deviation away from the minimum lambda there are 4 predictors.    
   
**Question 2**   
Looking the the predictors from the lambda = one standard deviation 0.0068 I see that the largest are X__7, X__5 and X__4. Meaning if I pick a lambda that is above the one standard deviation I will get these three.   
   
   
**Question 3**  
Seeing that the lambda that is one standard deviation away from the minimum lambda returns 4 predictors I know that I will have to pick a lambda that is larger than 0.0068. Meaning I will be outside the two lines on the plot above.   

```{r}
my_coef <- coef(glmmod, s=.0079)
print(my_coef)
```   
Picking a lambda of .0079 returns three predictors X__7, X__5 and X__4.
   
**Question 4**
```{r}
t3 <- data.frame(name = my_coef@Dimnames[[1]][my_coef@i + 1], coefficient = my_coef@x)
print(t3)
```   
The coefficient that are returned at lambda equil to .0079 are X__4 is 2.71673979, X__5 is -2.45651480, and X__7 is -0.03262433   






