---
title: "Jeff Kropelnicki"
name: "Homework 3"
date: "`r format(Sys.Date())`"
output: github_document
---   
   
For this weeks homework my job is to find the best two predictors from the patients data from last week. I need to use the lasso method. Below is full code to get the best two predictors from a regression model of 7.    
```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(knitr)
library(dplyr)
library(tidyr)
library(ggplot2)
```   
   
```{r, message=FALSE, warning=FALSE}
patients <- read.csv("patients.csv")

norm  <- patients %>% select(Age, Height, Weight)
norm <- as.data.frame(scale(norm))
#patients$Smoker <- as.factor(patients$Smoker)

norm_patients <- patients[,-c(1,3,4,6,7,8,10)]

dummy <- patients %>% select(Gender, Location, SelfAssessedHealthStatus, Smoker)
dummy <- data.frame(model.matrix(~., data=dummy)[,-1])

norm_patients <- cbind(norm_patients, norm, dummy) %>%
                select(Systolic,
                       Age,
                       Height,
                       Weight,
                       Smoker,
                       male = Gender.Male., 
                       Marys = Location.St..Mary.s.Medical.Center., 
                       VA = Location.VA.Hospital., 
                       fair = SelfAssessedHealthStatus.Fair., 
                       good = SelfAssessedHealthStatus.Good., 
                       poor = SelfAssessedHealthStatus.Poor.)


t1 <- lm(Systolic ~ Age + Weight + Height + male + Marys + VA + fair + good + poor + Smoker, data = norm_patients)
#summary(t1)
```   
   
      
```{r, message=FALSE, warning=FALSE}
library(glmnet)

# I need to breakup the predictors and responce make new matrix for X and Y. 
x <- as.matrix(norm_patients[,-1]) 
y <- as.double(as.matrix(norm_patients[, 1]))


set.seed(999)
cv.lasso <- cv.glmnet(x, y, alpha=1, nfolds = 10)

plot(cv.lasso)
plot(cv.lasso$glmnet.fit, xvar="lambda", label = TRUE)

```   
   
  
   
Lets see what the lambdas are for the min and 1 standard deviation from the min are    
```{r}
#The min 
t1 <- cv.lasso$lambda.min

#1 SD from the min
t2 <- cv.lasso$lambda.1se
```   
The minimum lambda is `r t1` and the lambda that is one standard deviation from the minimum is `r t2`
   

Lets see what predictors are selected with the minimum lambda         
```{r}
#predictors at mim lambda
coef(cv.lasso, s=cv.lasso$lambda.min)
```    
   
If I used the minimum lambda I would use 5 predictors, Age, Height, smoker, VA hospital and fair health.
    
       
Lets see what predictors are selected with lambda this is one standard deviation from the minimum.  
```{r}
#predictors at 1 sd from min lambda
coef(cv.lasso, s=cv.lasso$lambda.1se)
```   
   
If I used the lambda that is one standard deviation from the minimum I would use one predictor, Smoker.

**Question 5**   
Looking the the predictors from the lambda = minimum 0.3486 I see that the largest are smoker and Fair health. Meaning if I pick a lambda that is above the minimum I will get these two.    

**Question 6**    
Knowing the min lambda and the 1 SD lambda I know that I need to pick a lambda that is between those. I changed the number some and found that I can use lambda = 1 and I get the two predictors    
```{r}
my_coef <- coef(cv.lasso, s=1)
print(my_coef)
```   
With lambda equal to one I get two predictors smoker and fair health status.
    
            
**Question 7**   
```{r}
t3 <- data.frame(name = my_coef@Dimnames[[1]][my_coef@i + 1], coefficient = my_coef@x)
print(t3)
```   
   
The coefficient for smoker is 7.847697 and the coefficient for fair health status is -0.068012.

   























