---
title: "Homework 7"
author: "Jeff Kropelnicki"
date: '`r format(Sys.Date())`'
output:
  html_document: default
  pdf_document: default
---   
   
      
```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(knitr)
library(dplyr)
library(kableExtra)
library(readxl)
library(gplots)
library (ROCR)
library(caret)
library(nnet)
library(neuralnet)
```   
   
```{r, message=FALSE, warning=FALSE, echo=FALSE}
#Read in data.
data <- read_excel("CellDNA.xls", col_names = FALSE)
```   
   
      
Standerize the cell data and make the last column a 0 or 1. If a number is greater than 0 make it a 1.    
```{r}
#Make any number is the last column a 0 or 1
data$X__14[ data$X__14 > 0 ] = 1

#Normalize the data is the X columns 1-13
norm <- data[,1:13]
x <- as.matrix(scale(norm))
y <- as.matrix(data[,14])

#Take a look at the data
data <- as.data.frame(cbind(x,y))

kable(head(data), "html") %>%
          kable_styling(bootstrap_options = "striped", font_size = 10)


#Create the test and train data set here I use 70% for the training set
set.seed(777)
train1<-createDataPartition(data$X__14,p=0.7,list=FALSE)
train<-data[train1,]
test<-data[-train1,]
train$X__14 <- as.factor(train$X__14)
test$X__14 <- as.factor(test$X__14)


train$Y = class.ind(train$X__14)
train$X__14=NULL
```    
   
      
```{r}
#Create the first neural network on the train data
train.nnet <- nnet(Y~.,train, size=20, rang=0.07, Hess=FALSE, decay=15e-4, maxit=250, softmax = T, entropy=TRUE)

#Use test data to predict 
set.seed(777)
test.nnet<-predict(train.nnet,test,type=("class"))
#Print the table
table(test$X__14, test.nnet)

confusionMatrix(as.factor(test.nnet), test$X__14, mode = "prec_recall", positive="0")
confusionMatrix(as.factor(test.nnet), test$X__14, mode = "prec_recall", positive="1")
```   
Class 0   
Accuracy = 0.8986301  
Precision = 0.9455782    
Recall = 0.9297659   
   
Class 1   
Accuracy = 0.8986301   
Precision = 0.7042254          
Recall = 0.7575758  
   
Here I use a different r package because I want to see a plot of the neural net and get the ROC curve.      
```{r}
#This neural net needs data in a matrix
m <- model.matrix(  ~ Y + X__1 + X__2 + X__3 + X__4 + X__5 + 
                      X__6 + X__7 + X__8 + X__9 + X__10 + X__11 + X__12 + X__13, data=train)


#Make a second neural net with the matrix training data
nn <- neuralnet(Y1 ~ X__1 + X__2 + X__3 + X__4 + X__5 + X__6 + X__7 + 
                X__8 + X__9 + X__10 + X__11 + X__12 + X__13, 
                data=m, hidden=c(5,2), linear.output=FALSE, threshold=0.01)

#Plot the neural net
plot(nn)


#Subset data for a new test set for prediction
temp_test <- subset(test, select = c("X__1", "X__2", "X__3", "X__4", "X__5", "X__6", "X__7", 
                                     "X__8", "X__9", "X__10", "X__11", "X__12", "X__13"))

#Use test data to predict 
nn.results <- compute(nn, temp_test)
nn.results <- nn.results$net.result

detach(package:neuralnet,unload = T)
```   
   
 
 
This is code to get ROC curve I could not find a way to get both ROC curves in R. Over the last 2 homeworks there is no talk about 2 on the internet that I can find.       
```{r}
nn.pred = prediction(nn.results, test$X__14)
pref <- performance(nn.pred, "tpr", "fpr")
plot(pref)
```













